---
title: "RepData_Assignment1"
author: "KBen"
date: "Sunday, June 14, 2015"
output: html_document
---

## Loading and preprocessing the data
ind contains the indexes for omitting NA steps

```{r, echo=TRUE}
data <- read.csv("activity.csv")
ind = !is.na(data$steps)
```

## What is mean total number of steps taken per day?
```{r, echo=TRUE}
num.steps.day <- with(data[ind,] ,tapply(steps,date,sum))
     
library(ggplot2)
qplot(num.steps.day, geom="histogram",xlab="steps per day",main="Histogram of total number of steps with omitted NA values")
     
mean(num.steps.day,na.rm=T )
median(num.steps.day,na.rm=T )

```

## What is the average daily activity pattern?
```{r, echo=TRUE}
mean.interval <- with(data[ind,] ,tapply(steps,interval,mean))
plot(mean.interval,type="l")

mean.interval[which.max(mean.interval)]
```

## Imputing missing values
The number of NA values
```{r, echo=TRUE}
sum(!ind)
```

A strategy for filling in all of the missing values in the dataset. Using the mean for the 5-minute interval
```{r, echo=TRUE}
data.f <- data
data.f$steps[!ind] <- mean.interval[as.character(data.f$interval[!ind])]

num.steps.day.f <- with(data.f ,tapply(steps,date,sum))

qplot(num.steps.day.f, geom="histogram",xlab="steps per day",main="Histogram of total number of steps with imputed NA values")

mean(num.steps.day.f,na.rm=T )
median(num.steps.day.f,na.rm=T )
```

The mean and median are now the same. The mean is the same as with the data with NA omitted.

##Are there differences in activity patterns between weekdays and weekends?
```{r, echo=TRUE}
data.f$w.day = "weekday"
data.f$w.day[weekdays(as.Date(data.f$date)) %in% c('laugardagur','sunnudagur')] = "weekend"
data.f$w.day = as.factor(data.f$w.day)

library(lattice)
xyplot(steps~interval|w.day,data=data.f[order(data.f$interval),],type="l",horizontal=T)
```

People start activities later in the weekends
